---
title: "Land use land cover change due to energy transition-induced resource 
        extraction in Cabo Delgado, Mozambique"
author: "Emilinah Namaganda"
date: "2023-03-31"
output:
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Loading the necessary libraries

```{r libraries}
library(sf) 
library(lintr) 
library(rpart) 
library(dismo) 
library(ggplot2)
library(reshape2)
library(randomForest)
library(RStoolbox)
library(caret)
library(e1071)
library(rgdal)
library(terra)
library(sperrorest)
```

## Loading the necessary data

### Boundaries of the projects and their surrounding areas

The resource extracting (mining) projects (Twigg, GK and Total) are mainly 
hosted in three administrative posts (Balama, Ancuabe and Palma). 

Therefore, the areas of interest (aoi) include: 1) the administrative posts, 
2) the mine areas (as officially defined in the mining licences), and 3) the 
mine areas and their immediate surroundings, that is, including the mine/project 
affected communities (PACs) (I refer to this wider area as the 'study area'). 

I acquired the relevant shapefiles from various sources including:-
- Eduardo Mondlane University and the Ministry of Transport and Communication in 
  Mozambique, specifically the Spatial Development Program (SDP), for data on 
  the Administrative posts and location of communities. Similar shapefiles can 
  be acquired from online open access sources such as GRID: 
  https://data.grid3.org/
- https://portals.landfolio.com/mozambique/en/ for the boundaries of mines.

```{r}
knitr::opts_chunk$set(echo = FALSE)

sa_postsfn <- function(posts) {
  saposts           <- list.files(posts, 
                            pattern = "\\.shp$", 
                            full.names = TRUE)
  saposts_ls        <- list(saposts[grep("blma", saposts)],
                        saposts[grep("ancb", saposts)],
                        saposts[grep("plma", saposts)])
  saposts_ls        <- lapply(saposts_ls, vect)
  names(saposts_ls) <- c("blmapost", "ancbpost", "plmapost")
  return(saposts_ls)
}
sa_posts <- sa_postsfn("~/data/raw/mines/admposts")

sa_minesfn <- function(mines) {
  samines <- list.files(mines, 
                      pattern = "\\.shp$", 
                      full.names = TRUE)
  twigg   <- st_read(samines[3], 
                   query = "SELECT Mine FROM \"twigg\" WHERE FID = 0")
  gk      <- st_read(samines[1], 
                   query = "SELECT Mine FROM \"gk\" WHERE FID = 0")
  total   <- st_read(samines[2], 
                   query = "SELECT Name FROM \"total\" WHERE FID = 0")
  return(list("twigg" = twigg,
              "gk"    = gk,
              "total" = total))
}
sa_mines <- sa_minesfn("~/data/raw/mines/mines")

sa_pacsfn <- function(pacs) {
  PACs       <- list.files(pacs, 
                     pattern = "\\.shp$", 
                     full.names = TRUE)
  twigg_pacs <- st_read(PACs[3], 
                        query = "SELECT NomeN1 FROM \"twiggPACS\"")
  gk_pacs    <- st_read(PACs[1], 
                     query = "SELECT Community FROM \"gkPACS\"")
  total_pacs <- st_read(PACs[2], 
                        query = "SELECT Settlement FROM \"totalPACS\"")
  return(list("twigg_pacs" = twigg_pacs,
              "gk_pacs"    = gk_pacs,
              "total_pacs" = total_pacs))
}
sa_pacs <- sa_pacsfn("~/data/raw/mines/pacs")

study_areasfn <- function(mine_areas) {
  studyAreas          <- list.files(mine_areas, 
                            pattern = "\\.shp$", 
                            full.names = TRUE)
  studyAreasls        <- list(studyAreas[grep("blma", studyAreas)],
                              studyAreas[grep("ancb", studyAreas)],
                              studyAreas[grep("plma", studyAreas)])
  studyAreasls        <- lapply(studyAreasls, vect)
  names(studyAreasls) <- c("blma", "ancb", "plma")
  return(studyAreasls)
}
study_areas <- study_areasfn("~/data/raw/mines/stdyareas")
```

### Satellite images of the areas of interest

To study the change in land use and land cover (lulc) in the areas of interest, 
between 2015 to 2021, I use and therefore load bands 1 to 7 of the Landsat 8 
satellite. I also load the qa bands which indicate the pixels within the 
images that are covered by clouds.

All the images were downloaded from https://earthexplorer.usgs.gov/ 

```{r}
sa_bands17fn <- function(folders_images) {
  bands_all        <- list.files(folders_images,
                               recursive = T,
                               pattern = glob2rx("*B*"), 
                               full.names = TRUE)
  bands_17         <- bands_all[grep("B10", bands_all, invert = TRUE)]
  bands_17ls       <- list(bands_17[grep("2021.*165069", bands_17)],
                         bands_17[grep("2015.*165069", bands_17)],
                         bands_17[grep("2020.*165067", bands_17)],
                         bands_17[grep("2020.*164068", bands_17)],
                         bands_17[grep("2015.*165067", bands_17)],
                         bands_17[grep("2015.*164068", bands_17)],
                         bands_17[grep("2020.*165069", bands_17)],
                         bands_17[grep("2020.*164069", bands_17)],
                         bands_17[grep("2015.*165069", bands_17)],
                         bands_17[grep("2015.*164069", bands_17)])
  names(bands_17ls) <- (c("blma2021_b17",
                         "blma2015_b17",
                         "plma2020A_b17",
                         "plma2020B_b17",
                         "plma2015A_b17",
                         "plma2015B_b17",
                         "ancb2020A_b17",
                         "ancb2020B_b17",
                         "ancb2015A_b17",
                         "ancb2015B_b17"))
  bands             <- lapply(bands_17ls, rast)
  return(bands)
}
sa_bands <- sa_bands17fn("~/data/raw/images")

sa_qabandsfn <- function(folders_images) {
  qabands         <- list.files(folders_images,
                         recursive = T,
                         pattern = "\\PIXEL.TIF$",
                         full.names = TRUE)
  sa_qa           <- list(qabands[grep("2021", qabands)], 
                          qabands[grep("2015.*165069", qabands)],
                          qabands[grep("2020.*165067", qabands)],
                          qabands[grep("2020.*164068", qabands)],
                          qabands[grep("2015.*165067", qabands)],
                          qabands[grep("2015.*164068", qabands)],
                          qabands[grep("2020.*165069", qabands)],
                          qabands[grep("2020.*164069", qabands)],
                          qabands[grep("2015.*165069", qabands)],
                          qabands[grep("2015.*164069", qabands)])
  names(sa_qa)    <- list("blma2021_qa",
                          "blma2015_qa",
                          "plma2020A_qa",
                          "plma2020B_qa",
                          "plma2015A_qa",
                          "plma2015B_qa",
                          "ancb2020A_qa",
                          "ancb2020B_qa",
                          "ancb2015A_qa",
                          "ancb2015B_qa")
  sa_qarast <- lapply(sa_qa, rast)
  return(sa_qarast)
}
qa_bands <- sa_qabandsfn("~/data/raw/images")

qa_bandsCATfn <- function(qaBands) {
  qab            <- qaBands[[1]]
  activeCat(qab) <- 0
  return(qab)
}
qa_bandsfinal <- lapply(qa_bands, qa_bandsCATfn)
```

## Data preprocessing and analysis

### Preprocessing the images

To prepare the images for analysis, I create and use a cloud mask to mask out 
the clouded regions from the images. The cloud-free images are re-projected
into the right geographic coordinates for northern Mozambique and cropped to 
the levels of the administrative posts of interest.

In cases where two images are necessary to cover the relevant administrative 
post(s), the cropped images are mosaicked to create one image of each area for 
each year of study.

```{r}
moz_crs    <- "+init=EPSG:32737"

images_processfn  <- function(qaBands, bands17, aoi_post) {
  cloudmask     <- classify(qaBands, 
                            rbind(c(21824, 21824), 
                                  c(21952, 21952)),
                            others = NA)
  b17_clear     <- mask(bands17, cloudmask)
  b17_clearProj <- project(b17_clear, moz_crs)
  aoi_crop      <- crop(b17_clearProj, aoi_post, snap = "near", mask = TRUE)
  aoi_cropmsk   <- mask(aoi_crop, aoi_post)
  return(aoi_cropmsk)
}

aois_imagesfn <- function(aoi_fun, qaband, bands17, aoi_post) {
  aoisImages     <- list(
    "blma2021"   = aoi_fun(qaband[[1]], bands17[[1]], aoi_post[[1]]),
    "blma2015"   = aoi_fun(qaband[[2]], bands17[[2]], aoi_post[[1]]),
    "plma2020A"  = aoi_fun(qaband[[3]], bands17[[3]], aoi_post[[3]]),
    "plma2020B"  = aoi_fun(qaband[[4]], bands17[[4]], aoi_post[[3]]),
    "plma2015A"  = aoi_fun(qaband[[5]], bands17[[5]], aoi_post[[3]]),
    "plma2015B"  = aoi_fun(qaband[[6]], bands17[[6]], aoi_post[[3]]),
    "ancb2020A"  = aoi_fun(qaband[[7]], bands17[[7]], aoi_post[[2]]),
    "ancb2020B"  = aoi_fun(qaband[[8]], bands17[[8]], aoi_post[[2]]),
    "ancb2015A"  = aoi_fun(qaband[[9]], bands17[[9]], aoi_post[[2]]),
    "ancb2015B"  = aoi_fun(qaband[[10]], bands17[[10]], aoi_post[[2]]))
  AncbPlma2020_2015  <- list(
    "ancb2020"  = mosaic(aoisImages$ancb2020A, aoisImages$ancb2020B),
    "ancb2015"  = mosaic(aoisImages$ancb2015A, aoisImages$ancb2015B),
    "plma2020"  = mosaic(aoisImages$plma2020A, aoisImages$plma2020B),
    "plma2015"  = mosaic(aoisImages$plma2015A, aoisImages$plma2015B))
  aoisImages_all <- c(aoisImages, AncbPlma2020_2015)
  aois_Images          <- aoisImages_all[-c(3:10)]
  return(aois_Images)
}
aois_images <- aois_imagesfn(images_processfn, 
                             qa_bandsfinal, 
                             sa_bands, 
                             sa_posts)

aois_imagesfinal_fn <- function(aois_imagesls) {
  imageslist                      <- aois_imagesls
  for (i in 1:length(imageslist)) {
    names(imageslist[[i]][[1:7]]) <- gsub(pattern = ".*SR",
                                          replacement = names(imageslist[i]),
                                          names(imageslist[[i]][[1:7]]))
  }
  return(imageslist)
}
aois_imagesfinal <- aois_imagesfinal_fn(aois_images)
```

### Creating training samples for image classification

To classify the images of the areas of interest, I created training samples 
from RGB composites of the areas. To aid with distinguishing features, I  
utilized ndvi layers, Google Earth (https://earth.google.com/web/), knowledge of
the study areas from field visits, and published projects' environmental 
impact assessment reports:
 - Twigg (http://www.cesnet.co.za/pubdocs/Syrah%20Balama%20Graphite%20Mine%20
   Port%20CB290814/Parte%20II%20ESHIAR.pdf)
 - Total (https://mzlng.totalenergies.co.mz/sites/g/files/wompnd2311/f/atoms/
   files/mz-000-am1-hs-rpt-00002_eshia_executive_summary_and_update-05-19-2020.
   pdf).

I created approximately 500 pixels per image. This process was conducted in 
ArcGIS Pro Version 2.8.3, but it can also be done in open access QGIS 
(https://www.qgis.org/en/site/).

For improved reliability of the classification model, training and/or validation 
samples should be generated from groundtruth data, where possible. However, 
given the wide extent of the study areas, and the inaccessibility of some of 
these areas due to an armed conflict, training and validation 
data for this study were collected from the same image data, supplemented with 
the aforementioned secondary material. The generation of training samples from 
imagery as a surrogate for ground observation has been utilized in other studies 
(Foody, 2002).

```{r}
comp   <- function(bands17, x, y, z) {
  comps    <- c(bands17[[x]], bands17[[y]], bands17[[z]])
  return(comps)
}

aois_RGB <- function(bands17) {
  aois_comp        <- list("blma2021_b543"  = comp(bands17[[1]], 5, 4, 3),
                           "blma2015_b543"  = comp(bands17[[2]], 5, 4, 3),
                           "blma2021_b654"  = comp(bands17[[1]], 6, 5, 4),
                           "blma2015_b654"  = comp(bands17[[2]], 6, 5, 4),
                           "ancbe2020_b543" = comp(bands17[[3]], 5, 4, 3),
                           "ancbe2015_b543" = comp(bands17[[4]], 5, 4, 3),
                           "ancbe2020_b654" = comp(bands17[[3]], 6, 5, 4),
                           "ancbe2015_b654" = comp(bands17[[4]], 6, 5, 4),
                           "plma2020_b543"  = comp(bands17[[5]], 5, 4, 3),
                           "plma2015_b543"  = comp(bands17[[6]], 5, 4, 3),
                           "plma2020_b654"  = comp(bands17[[5]], 6, 5, 4),
                           "plma2015_b654"  = comp(bands17[[6]], 6, 5, 4))
  aois_compnames   <- paste0(names(aois_comp), ".tif")
  names(aois_comp) <- aois_compnames
  for (i in 1:length(aois_comp)) {
  aoisComp         <- writeRaster(aois_comp[[i]],
                 file.path("~/results/output", 
                           names(aois_comp[i])),
                 overwrite=TRUE,
                 filetype="GTiff")
  }
  return(aois_comp)
}
aois_compst <- aois_RGB(aois_imagesfinal)

indices <- function(bands17, x, y) {
  bandx   <- bands17[[x]]
  bandy   <- bands17[[y]]
  indices <- (bandx - bandy)/(bandx + bandy)
  return(indices)
}

ndvi_fun <- function(bands17) {
  ndvi         <- list("blma2021_ndvi" = indices(bands17[[1]], 5, 4),
                       "blma2015_ndvi" = indices(bands17[[2]], 5, 4),
                       "ancb2020_ndvi" = indices(bands17[[3]], 5, 4),
                       "ancb2015_ndvi" = indices(bands17[[4]], 5, 4),
                       "plma2020_ndvi" = indices(bands17[[5]], 5, 4),
                       "plma2015_ndvi" = indices(bands17[[6]], 5, 4))
  ndvinames    <- paste0(names(ndvi), ".tif")
  names(ndvi)  <- ndvinames
  for (i in 1:length(ndvi)) {
  writeRaster(ndvi[[i]],
                 file.path("~/results/output", names(ndvi[i])),
                 overwrite=TRUE,
                 filetype="GTiff")
  }
  return(ndvi)
}
aois_ndvi <- ndvi_fun(aois_imagesfinal)

train_polys <- list(
  "blma2021_train" = vect("~/data/temp/training-samples/blma2021_train500.shp"),
  "blma2015_train" = vect("~/data/temp/training-samples/blma2015_train500.shp"),
  "ancb2020_train" = vect("~/data/temp/training-samples/ancb2020_train500.shp"),
  "ancb2015_train" = vect("~/data/temp/training-samples/ancb2015_train500.shp"),
  "plma2020_train" = vect("~/data/temp/training-samples/plma2020_train500.shp"),
  "plma2015_train" = vect("~/data/temp/training-samples/plma2015_train500.shp"))
```

For the supervised classification, I follow the workflow suggested by 
Ghosh & Hijmans (2021). From the training sample polygons, I select sample 
points for use in the lulc classification process. The spectral profiles of the 
samples can be inspected to ensure adequate differentiation among the lulc 
classes. 

```{r}
samps <- function(train, image_aoi) {
                set.seed(555)
  Tsamps     <- spatSample(train, 1000, method = "regular")
                print(table(as.character(Tsamps$Classname)))
  sampsXY    <- as.matrix(geom(Tsamps)[,c('x','y')])
  sampsdf    <- extract(image_aoi[[3:6]], sampsXY)
  samps      <- data.frame(class = Tsamps$Classname, sampsdf)
  sampsGeo   <- cbind(samps, sampsXY)
  sampsGeoSF <- st_as_sf(sampsGeo, coords = c("x", "y"), crs = moz_crs)
  return(sampsGeoSF)
}

sa_samps        <- mapply(train_polys, 
                          aois_imagesfinal, 
                          FUN = samps, 
                          SIMPLIFY = F)
names(sa_samps) <- c("blma2021_samps",
                     "blma2015_samps",
                     "ancb2020_samps",
                     "ancb2015_samps",
                     "plma2020_samps",
                     "plma2015_samps")

sampsSP <- function(samps) {
  sampsSP             <- aggregate(st_drop_geometry(samps)[, -1], 
                                   list(samps$class), 
                                   mean)
  rownames(sampsSP)   <- sampsSP[, 1]
  sampsSP             <- sampsSP[, -1]
  names(sampsSP)[1:4] <- 1:4
  sampsSP             <- t(sampsSP)
  sampsSP             <- as.data.frame(sampsSP)
  sampsSP             <- cbind(bands = rownames(sampsSP), sampsSP)
  rownames(sampsSP)   <- NULL
  sampsSP             <- melt(sampsSP, 
                              id.vars = "bands", 
                              variable.name = "lulc")
  sampsSP$bands       <- as.numeric(sampsSP$bands)
  return(sampsSP)
}
allsampsSP <- lapply(sa_samps, sampsSP)

lulc_class <- list(BlmaAncb = c("built & bare",
                                "dense vegetation",
                                "sparse vegetation",
                                "water"),
                     Plma   = c("built & bare",
                                "dense vegetation",
                                "sparse vegetation",
                                "wetland"))

lulc_cols <- list(BlmaAncb = c('darkorange', 
                               'darkgreen', 
                               'green', 
                               'deepskyblue'),
                  Plma     = c('darkorange', 
                                'darkgreen', 
                                'green', 
                                'lightcyan'))

sampsSPgraph             <- list(
  "blma2021sp" = ggplot(allsampsSP[[1]], aes(bands, value)) +
    geom_line(aes(color = lulc), size = 1) +
    ggtitle("Spectral Profile Balama 2021") +
    labs(x = "Bands", y = "Reflectance") +
    scale_colour_manual("LULC", 
                    values = lulc_cols$BlmaAncb,
                    labels = lulc_class$BlmaAncb),
  "blma2015sp" = ggplot(allsampsSP[[2]], aes(bands, value)) +
    geom_line(aes(color = lulc), size = 1) +
    ggtitle("Spectral Profile Balama 2015") +
    labs(x = "Bands", y = "Reflectance") +
    scale_colour_manual("LULC", 
                    values = lulc_cols$BlmaAncb,
                    labels = lulc_class$BlmaAncb),
  "ancb2020sp" = ggplot(allsampsSP[[3]], aes(bands, value)) +
    geom_line(aes(color = lulc), size = 1) +
    ggtitle("Spectral Profile Ancuabe 2020") +
    labs(x = "Bands", y = "Reflectance") +
    scale_colour_manual("LULC", 
                    values = lulc_cols$BlmaAncb,
                    labels = lulc_class$BlmaAncb),
  "ancb2015sp" = ggplot(allsampsSP[[4]], aes(bands, value)) +
    geom_line(aes(color = lulc), size = 1) +
    ggtitle("Spectral Profile Ancuabe 2015") +
    labs(x = "Bands", y = "Reflectance") +
    scale_colour_manual("LULC", 
                    values = lulc_cols$BlmaAncb,
                    labels = lulc_class$BlmaAncb),
  "plma2020sp" = ggplot(allsampsSP[[5]], aes(bands, value)) +
    geom_line(aes(color = lulc), size = 1) +
    ggtitle("Spectral Profile Palma 2020") +
    labs(x = "Bands", y = "Reflectance") +
    scale_colour_manual("LULC", 
                    values = lulc_cols$Plma,
                    labels = lulc_class$Plma),
  "plma2015sp" = ggplot(allsampsSP[[6]], aes(bands, value)) +
    geom_line(aes(color = lulc), size = 1) +
    ggtitle("Spectral Profile Palma 2015") +
    labs(x = "Bands", y = "Reflectance") +
    scale_colour_manual("LULC", 
                    values = lulc_cols$Plma,
                    labels = lulc_class$Plma))
```

## Image classification 

I classify each image using the random forest model. Afterwards, I crop the 
output by the study and mine areas. 

```{r}
comp_4bs   <- function(bands17, w, x, y, z) {
  comps    <- c(bands17[[w]], bands17[[x]], bands17[[y]], bands17[[z]])
  return(comps)
}
aois_4bs <- lapply(aois_imagesfinal, w = 3, x = 4, y = 5, z = 6, comp_4bs)


lulc <- function(samps, aoi) {
                set.seed(555)
  model      <- randomForest(as.factor(class)~., 
                             data = st_drop_geometry(samps),
                             method = 'class', 
                             minsplit = 5)
  classified <- predict(aoi, model, na.rm = TRUE)
  return(classified)
}
aois_lulc <- mapply(lulc, sa_samps, aois_4bs)


sa_lulcfn <- function(lulc_ls) {
  LULC       <- list("blma2021_LULC" = crop(lulc_ls[[1]], study_areas[[1]]),
                     "blma2015_LULC" = crop(lulc_ls[[2]], study_areas[[1]]),
                     "ancb2021_LULC" = crop(lulc_ls[[3]], study_areas[[2]]),
                     "ancb2015_LULC" = crop(lulc_ls[[4]], study_areas[[2]]),
                     "plma2020_LULC" = crop(lulc_ls[[5]], study_areas[[3]]),
                     "plma2015_LULC" = crop(lulc_ls[[6]], study_areas[[3]]))
  LULCnames   <- paste0(names(LULC), ".tif")
  names(LULC) <- LULCnames
  for (i in 1:length(LULC)) {
  writeRaster(LULC[[i]],
                 file.path("~/results/output", names(LULC[i])),
                 overwrite=TRUE,
                 filetype="GTiff")
  }
  return(LULC)
}
sa_LULC <- sa_lulcfn(aois_lulc)


lulcMines <- list()

lulc_minesfn <-  function(aoiLULC, mine_area) {
  blma2021_twigg  <- crop(aoiLULC[[1]], mine_area[[1]])
  blma2015_twigg  <- crop(aoiLULC[[2]], mine_area[[1]])
  ancb2021_gk     <- crop(aoiLULC[[3]], mine_area[[2]])
  ancb2015_gk     <- crop(aoiLULC[[4]], mine_area[[2]])
  plma2021_total  <- crop(aoiLULC[[5]], mine_area[[3]])
  plma2021_total  <- mask(plma2021_total, mine_area[[3]])
  plma2015_total  <- crop(aoiLULC[[6]], mine_area[[3]])
  plma2015_total  <- mask(plma2015_total, mine_area[[3]])
  lulcMines       <- list("blma2021_twigg" = blma2021_twigg,
                         "blma2015_twigg"  = blma2015_twigg,
                         "ancb2021_gk"     = ancb2021_gk,
                         "ancb2015_gk"     = ancb2015_gk,
                         "plma2020_total"  = plma2021_total,
                         "plma2015_total"  = plma2015_total)
  return(lulcMines)
}
lulc_mines <- lulc_minesfn(aois_lulc, sa_mines)
```

#### Plots of study area land use land cover 

```{r}
lulcdf_blmaAncb <- data.frame(id = 1:4, lulc = lulc_class$BlmaAncb)
lulcdf_plma     <- data.frame(id = 1:4, lulc = lulc_class$Plma)

saLULCdf <- sa_LULC

levels(saLULCdf[[1]]) <- lulcdf_blmaAncb
levels(saLULCdf[[2]]) <- lulcdf_blmaAncb
levels(saLULCdf[[3]]) <- lulcdf_blmaAncb
levels(saLULCdf[[4]]) <- lulcdf_blmaAncb
levels(saLULCdf[[5]]) <- lulcdf_plma
levels(saLULCdf[[6]]) <- lulcdf_plma


for (i in 1:length(saLULCdf)) {
  saLULCdf[[i]] <- as.data.frame(saLULCdf[[i]], xy=TRUE)
}


ggplot() +
  geom_tile(data = saLULCdf[[1]], aes(x = x, y = y, fill = lulc)) +
  geom_sf(data   = sa_mines[[1]], colour = "black", fill = NA, size = 1) +
  geom_sf(data   = sa_pacs[[1]], fill = NA) +
  geom_sf_text(data = sa_pacs[[1]], aes(label = NomeN1)) +
  labs(x = NULL, 
       y = NULL, 
       title        = "Balama 2021 land use land cover") +
  scale_fill_manual("LULC",
                    na.translate = FALSE,
                    values = lulc_cols[[1]],
                    labels = lulc_class[[1]]) +
  coord_sf(datum    = st_crs(sa_mines[[1]]))

ggplot() +
  geom_tile(data = saLULCdf[[2]], aes(x = x, y = y, fill = lulc)) +
  geom_sf(data   = sa_mines[[1]], colour = "black", fill = NA, size = 1) +
  geom_sf(data   = sa_pacs[[1]], fill = NA) +
  geom_sf_text(data = sa_pacs[[1]], aes(label = NomeN1)) +
  labs(x = NULL, 
       y = NULL, 
       title        = "Balama 2015 land use land cover") +
  scale_fill_manual("LULC",
                    na.translate = FALSE,
                    values = lulc_cols[[1]],
                    labels = lulc_class[[1]]) +
  coord_sf(datum   = st_crs(sa_mines[[1]]))

ggplot() +
  geom_tile(data = saLULCdf[[3]], aes(x = x, y = y, fill = lulc)) +
  geom_sf(data   = sa_mines[[2]], colour = "black", fill = NA, size = 1) +
  geom_sf(data   = sa_pacs[[2]], fill = NA) +
  geom_sf_text(data = sa_pacs[[2]], aes(label = Community)) +
  labs(x = NULL, 
       y = NULL, 
       title       = "Ancuabe 2020 land use land cover") +
  scale_fill_manual("LULC",
                    na.translate = FALSE,
                    values = lulc_cols[[1]],
                    labels = lulc_class[[1]]) +
  coord_sf(datum   = st_crs(sa_mines[[2]]))

ggplot() +
  geom_tile(data = saLULCdf[[4]], aes(x = x, y = y, fill = lulc)) +
  geom_sf(data   = sa_mines[[2]], colour = "black", fill = NA, size = 1) +
  geom_sf(data   = sa_pacs[[2]], fill = NA) +
  geom_sf_text(data = sa_pacs[[2]], aes(label = Community)) +
  labs(x = NULL, 
       y = NULL, 
       title       = "Ancuabe 2015 land use land cover") +
  scale_fill_manual("LULC",
                    na.translate = FALSE,
                    values = lulc_cols[[2]],
                    labels = lulc_class[[2]]) +
  coord_sf(datum  = st_crs(sa_mines[[2]]))

ggplot() +
  geom_tile(data = saLULCdf[[5]], aes(x = x, y = y, fill = lulc)) +
  geom_sf(data   = sa_mines[[3]], colour = "black", fill = NA, size = 1) +
  geom_sf(data   = sa_pacs[[3]], fill = NA) +
  geom_sf_text(data = sa_pacs[[3]], aes(label = Settlement)) +
  labs(x = NULL, 
       y = NULL, 
       title        = "Palma 2020 land use land cover") +
  scale_fill_manual("LULC",
                    na.translate = FALSE,
                    values = lulc_cols[[2]],
                    labels = lulc_class[[2]]) +
  coord_sf(datum   = st_crs(sa_mines[[3]]))

ggplot() +
  geom_tile(data = saLULCdf[[6]], aes(x = x, y = y, fill = lulc)) +
  geom_sf(data   = sa_mines[[3]], colour = "black", fill = NA, size = 1) +
  geom_sf(data   = sa_pacs[[3]], fill = NA) +
  geom_sf_text(data = sa_pacs[[3]], aes(label = Settlement)) +
  labs(x = NULL, 
       y = NULL, 
       title       = "Palma 2015 land use land cover") +
  scale_fill_manual("LULC",
                    na.translate = FALSE,
                    values = lulc_cols[[2]],
                    labels = lulc_class[[2]]) +
  coord_sf(datum   = st_crs(sa_mines[[3]]))
```

## Change detection

The classified images of the study and mine areas include a 'count' column 
which includes area in pixels taken up by each lulc class. I multiply these 
areas by 0.0009 to convert them to square kilometers. I use the changes in 
these lulc areas for each location over the two time stamps (2021/2020 & 2015) 
to determine the change in lulc due to the three energy transition-related 
resource extraction areas.


```{r}
for (i in 1:length(sa_LULC)) {
  lulc_freq <- freq(sa_LULC[[i]])*0.0009
  print(lulc_freq)
}

for (i in 1:length(lulc_mines)) {
  lulc_freqMines <- freq(lulc_mines[[i]])*0.0009
  print(lulc_freqMines)
}
```


## Accuracy assessment

I assess the accuracy of the lulc classification model using 
k-fold cross validation. I also conduct a spatial cross validation to assess the 
accuracy of the model, while taking into account possible spatial 
autocorrelation (Brenning, 2012), which is common with spatial data. The 
water and wetland classes are spatially-small, compared to the rest of the 
classes. This imbalance likely contributes to an optimistic assessment of the 
accuracy of the classification. I repeat the spatial cross validation, excluding
the water and wetland classes.

```{r}
accuracies <- list()

lulc_accsFN <- function(sampsRF) {
  for (k in 1:5) {
                  set.seed(555)
    sampsRF         <- st_drop_geometry(sampsRF)
    sampGrps        <- kfold(sampsRF, k = 5, by = sampsRF$class)
                       print(table(sampGrps))
    train           <- sampsRF[sampGrps!= k, ]
    test            <- sampsRF[sampGrps == k, ]
    model           <- randomForest(as.factor(class)~., 
                                    data=train, 
                                    method = 'class', 
                                    minsplit = 5)
    classfd         <- predict(model, test, na.rm = TRUE)
    accuracies[[k]] <- cbind(test$class, as.integer(classfd))
  }
  return(accuracies)
}
lulc_accs <-  lapply(sa_samps, lulc_accsFN)


lulc_acc <- function(accuracies, lulcdf) {
  acctidy            <- do.call(rbind, accuracies)
  acctidy            <- data.frame(acctidy)
  colnames(acctidy)  <- c('observed', 'predicted')
  confmat            <- table(acctidy)
  colnames(confmat)  <- lulcdf$class
  testSamps          <- sum(confmat)
  testSamps_true     <- diag(confmat)
  overallAcc         <- sum(testSamps_true)/testSamps
  obsCases           <- apply(confmat, 1, sum)
  obsCases_true      <- obsCases/testSamps
  predCases          <- apply(confmat, 2, sum)
  predCases_true     <- predCases/testSamps
  expAcc             <- sum(obsCases_true*predCases_true)
  kappa              <- (overallAcc - expAcc)/(1 - expAcc)
  ProdAcc            <- testSamps_true/predCases
  UserAcc            <- testSamps_true/obsCases
  outAcc             <- data.frame(ProdAcc, UserAcc)
  lulc_acc           <- list(confmat, overallAcc, kappa, outAcc)
  return(lulc_acc)
}

LULC_accsBlmaAncb <- lapply(lulc_accs[1:4], lulcdf_blmaAncb, FUN=lulc_acc)
LULC_accsPlma     <- lapply(lulc_accs[5:6], lulcdf_blmaAncb, FUN=lulc_acc)


samps_val <- function(train, image_aoi) {
                set.seed(555)
  Tsamps     <- spatSample(train, 1000, method = "regular")
  sampsXY    <- as.matrix(geom(Tsamps)[,c('x','y')])
  sampsdf    <- extract(image_aoi[[3:6]], sampsXY)
  samps      <- data.frame(class = as.factor(Tsamps$Classname), sampsdf)
  sampsGeo   <- cbind(samps, sampsXY)
  return(sampsGeo)
}
sa_samps_val        <- mapply(train_polys, 
                          aois_imagesfinal, 
                          FUN = samps_val, 
                          SIMPLIFY = F)
names(sa_samps_val) <- c("blma2021_sampsV",
                         "blma2015_sampsV",
                         "ancb2020_sampsV",
                         "ancb2015_sampsV",
                         "plma2020_sampsV",
                         "plma2015_sampsV")

blma21 <- class~blma2021_B3 + blma2021_B4 + blma2021_B5 + blma2021_B6
blma15 <- class~blma2015_B3 + blma2015_B4 + blma2015_B5 + blma2015_B6
ancb20 <- class~ancb2020_B3 + ancb2020_B4 + ancb2020_B5 + ancb2020_B6
ancb15 <- class~ancb2015_B3 + ancb2015_B4 + ancb2015_B5 + ancb2015_B6
plma20 <- class~plma2020_B3 + plma2020_B4 + plma2020_B5 + plma2020_B6
plma15 <- class~plma2015_B3 + plma2015_B4 + plma2015_B5 + plma2015_B6

forms  <- c(blma21, blma15, ancb20, ancb15, plma20, plma15)
lapply(forms, as.formula)

mypred.rf <- function(object, newdata) {
  predict(object, newdata, na.rm = TRUE)
}

spCV <- function(valsamps, form){
              set.seed(555)
  spCV_all <- sperrorest(data = valsamps, 
                     formula = form,
                     model_fun = randomForest,
                     model_args = list(method = 'class'),
                     pred_fun = mypred.rf,
                     smp_fun = partition_kmeans,
                     smp_args = list(repetition = 1:10, nfold = 5))
  return(spCV_all)
}

blma21_accV <- spCV(sa_samps_val[[1]], forms[[1]])
blma15_accV <- spCV(sa_samps_val[[2]], forms[[2]])
ancb20_accV <- spCV(sa_samps_val[[3]], forms[[3]])
ancb15_accV <- spCV(sa_samps_val[[4]], forms[[4]])
plma20_accV <- spCV(sa_samps_val[[5]], forms[[5]])
plma15_accV <- spCV(sa_samps_val[[6]], forms[[6]])


train_polys_3cls <- train_polys

tply_3cls <- list()

for(i in 1:length(train_polys_3cls)) {
  cls3 <- train_polys_3cls[[i]][!((train_polys_3cls[[i]][[1]]=="water") | (train_polys_3cls[[i]][[1]]=="wetland")), ]
  tply_3cls[[i]]   <- cls3
}

names(tply_3cls) <- names(train_polys_3cls)

samps_val_3cls <- function(train, image_aoi) {
  set.seed(555)
  Tsamps     <- spatSample(train, 1000, method = "random")
  sampsXY    <- as.matrix(geom(Tsamps)[,c('x','y')])
  sampsdf    <- extract(image_aoi[[3:6]], sampsXY)
  samps      <- data.frame(class = as.factor(Tsamps$Classname), sampsdf)
  sampsGeo   <- cbind(samps, sampsXY)
  return(sampsGeo)
}

sa_samps_val_3cls        <- mapply(tply_3cls, 
                                   aois_imagesfinal, 
                                   FUN = samps_val_3cls, 
                                   SIMPLIFY = F)
names(sa_samps_val_3cls) <- c("blma2021_sampsV",
                              "blma2015_sampsV",
                              "ancb2020_sampsV",
                              "ancb2015_sampsV",
                              "plma2020_sampsV",
                              "plma2015_sampsV")

blma21_accV_3cls <- spCV(sa_samps_val_3cls[[1]], forms[[1]])
blma15_accV_3cls <- spCV(sa_samps_val_3cls[[2]], forms[[2]])
ancb20_accV_3cls <- spCV(sa_samps_val_3cls[[3]], forms[[3]])
ancb15_accV_3cls <- spCV(sa_samps_val_3cls[[4]], forms[[4]])
plma20_accV_3cls <- spCV(sa_samps_val_3cls[[5]], forms[[5]])
plma15_accV_3cls <- spCV(sa_samps_val_3cls[[6]], forms[[6]])

summary(ancb15_accV_3cls$error_rep$test_error)
summary(ancb15_accV_3cls$error_rep$test_accuracy)
```

## References

Brenning, A., 2012. Spatial cross-validation and bootstrap for the assessment 
of prediction rules in remote sensing: The R package sperrorest. In 2012 IEEE 
international geoscience and remote sensing symposium (5372-5375). IEEE. 

Foody, G. M., 2002. Status of land cover classification accuracy assessment. 
Remote sensing of environment, 80(1), 185-201.

Ghosh, A., & Hijmans, R. J., 2021. Remote sensing image analysis with R. 
Rspatial.org.

