---
title: "Land use land cover change due to green extractivism in Mozambique"
author: "Emilinah Namaganda"
date: "7/18/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

Using Rmarkdown, open access techniques, and remote sensing data is useful for
monitoring land use land cover change in general in low income areas. It is also
useful in areas where corruption is rife or where transparency in the extractive
industry is rare as anyone can double-check reports from government, companies, 
or any other stakeholders. A project to monitor activities can be opened up
online, verified, and updated by anyone.

Geo-politically, open access platforms also allow Africans or people in other low
income countries to use these platforms to develop their own resources that are
best for their needs. E.g. Gitlab a hosting site for open access project is itself
hosted on a free and open source platform.

If done, earlier in the project, it is
a useful baseline and change tool. This is one of the benefits of this study 
which is being conducted at the beginning of natural gas and graphite extraction 
activities in Cabo Delgado, Mozambique.

With google earth engine (GEE) you need a stable internet connection. With R, 
you can make local copies of online repositories for later local analysis. 
This may be helpful for analysis in areas with scanty internet connectivity.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the necessary libraries

```{r }

library(terra) # raster data
library(sf) # vector data
library("lintr")

```

## Preprocessing Balama images 2021

Users without administrative boundary data can use open access data from the 
GRID3 DATA HUB: https://data.grid3.org/

Mine areas can be constructed from geolocations from the Mozambique Mining
Cadastre Map Portal: https://portals.landfolio.com/mozambique/en/

Landsat image data were downloaded from the USGS website:
https://earthexplorer.usgs.gov/

```{r}

# loading relevant image bands -----------


bands17 <- function(images_path) {
  bands17 <- list.files(images_path,
    pattern = glob2rx("*B*"),
    full.names = TRUE
  )
  bands17 <- rast(bands17[1:7])
  return(bands17)
}

blma2021_b17 <- bands17("~/data/raw/images/2021/0609")
blma2021_b17


# removing cloud cover pixels, re-projecting to correct CRS, crop to AOI ------

moz_crs <- "+init=EPSG:32737"
blma <- vect(
  x = "~/data/raw/balama/blma.shp",
  query = "SELECT POSTO FROM \"blma\""
)
syrah <- vect(
  x = "~/data/raw/balama/syrah.shp",
  query = "SELECT Mine FROM \"syrah\" WHERE FID = 0"
)
syrah_buf <- buffer(syrah, 9000)


blma2021_qa <- rast(
  "~/data/raw/images/2021/0609/LC08_L2SP_165069_20210609_20210615_02_T1_QA_PIXEL.TIF"
)
activeCat(blma2021_qa) <- 0
plot(blma2021_qa)


aoi <- function(QA_band, bands_17, mine_area) {
  cloudsmask <- classify(QA_band, rbind(c(21824, 21824), c(21952, 21952)),
    othersNA = TRUE
  )
  b17_clear <- mask(bands17, cloudsmask)
  b17_clear_proj <- project(b17_clear, moz_crs)
  aoi_mine_area <- crop(b17_clear_proj, mine_area, snap = "near", mask = TRUE)
  return(aoi_mine_area)
}

blma2021 <- aoi(blma2021_qa, blma2021_b17, syrah_buf)
names(blma2021) <- gsub(
  pattern = "LC08_L2SP_165069_20210609_20210615_02_T1_SR",
  replacement = "blma2021",
  names(blma2021))

plot(blma2021$blma2021_B4)
plot(syrah, add = TRUE)

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent
printing of the R code that generated the plot.
